library(tidyverse)
activity <- read.csv("activity.csv")


activity$date<- as.Date.factor(activity$date,"%m/%d/%Y")

weekday<- weekdays(activity$date)
activity <-cbind(activity, weekday)
summary(activity)

    steps             date          interval        weekday         
 Min.   :  0.00   Min.   :NA      Min.   :   0.0   Length:17568      
 1st Qu.:  0.00   1st Qu.:NA      1st Qu.: 588.8   Class :character  
 Median :  0.00   Median :NA      Median :1177.5   Mode  :character  
 Mean   : 37.38   Mean   :NA      Mean   :1177.5                     
 3rd Qu.: 12.00   3rd Qu.:NA      3rd Qu.:1766.2                     
 Max.   :806.00   Max.   :NA      Max.   :2355.0                     
 NA's   :2304     NA's   :17568

## 1. What is mean total number of steps taken per day?
```{r echo=TRUE}
activity.tsteps<- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))

names(activity.tsteps)<- c("dates", "steps")

hist(activity.tsteps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

## Mean number of steps taken per day
mean(activity.tsteps$steps)
## [1] 9354.23

## Median number of steps taken per day
median(activity.tsteps$steps)
## [1] 10395

## 2. What is the average daily activity pattern?
```{r echo=TRUE}
average.daily.activity<- aggregate(activity$steps, by= list(activity$interval), FUN = mean , na.rm = TRUE)

names(average.daily.activity)<-c("interval", "mean")

plot(average.daily.activity$interval, average.daily.activity$mean, type = "l", xlab = "Interval", ylab = "Average number of steps", main = "Average number of steps per interval")
```

## 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

average.daily.activity[which.max(average.daily.activity$mean),]$interval
##  [1] 835

## 3. Imputing missing values

## Calculate and report the total number of missing values in the dataset 

sum(is.na(activity$steps))

## [1] 2304


## Devise a strategy for filling in all of the missing values in the dataset.

clean.steps<- average.daily.activity$mean[match(activity$interval,average.daily.activity$interval)]
head(clean.steps)
## [1] 1.7169811 0.3396226 0.1320755 0.1509434 0.0754717 2.0943396

## Create a new dataset that is equal to the original dataset but with the missing data filled in.

activity.clean <- transform(activity, steps = ifelse(is.na(activity$steps), yes = clean.steps, no = activity$steps))

total.clean.steps<- aggregate(steps ~ date, activity.clean, sum)

names(total.clean.steps)<- c("date", "daily.steps")

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

hist(total.clean.steps$daily.steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))

## Mean of the total number of steps taken per day

mean(total.clean.steps$daily.steps)
## [1] 10766.19

## Median of the total number of steps taken per day

median(total.clean.steps$daily.steps)

##[1] 10766.19

## 4. Are there differences in activity patterns between weekdays and weekends?

## Create a new factor variable in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        }
        
 
## A panel plot containing a time series plot (type=“l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

activity.datetype <- aggregate(steps~interval+datetype, activity,mean, na.rm =TRUE)
ggplot(activity.datetype, aes(x = interval, y = steps, color = datetype))+ geom_line() + labs(title = "Average daily steps by datetype", x = "Interval", y = "Average number of steps") + facet_wrap(~datetype, ncol = 1, nrow = 2) 


